---
title: "SNL Speakers"
author: "Jonathan Peelle"
#date: "October 19, 2015"
output: pdf_document
---


Introduction and motivation
---------------------------


I've uploaded the data and analysis (i.e., this document, written in Rmarkdown) to a public github repository so that others can contribute as well: <https://github.com/jpeelle/SNLspeakers>.



Data
----

I obtained data on speakers from previous PDFs of meeting programs, downloaded from <http://www.neurolang.org/previous/> on October 19, 2015.

I put together a spreadsheet listing speakers for each year, along with their gender, country, type of talk (keynote, debate, symposium), and so on, in `SNLspeakerdata.tsv` (tab-separated text file). The data cover 7 years, 2009-2015 inclusive.

I determined sex and country information based on my best guesses from the program and assumptions based on photos or first names. For "board membership" I judged based on the current board published in the PDF program from each year. (My impression is that I may have missed some speakers who had society affiliations of other sorts.)

I'm including all relevant code in this document, which you should be able to paste in to an R console and recreate the numbers and plots. You can see the Rmarkdown source (`SNLspeakers.Rmd`) for the plain-text version.


Results
-------

Between 2009-2015 SNL has had XX total speakers, of whom XX (XX%) have been women.



Comment
-------



#Acknowledgments
#---------------


You can embed an R code chunk like this:

```{r}
summary(cars)
```

You can also embed plots, for example:

```{r, echo=FALSE}
plot(cars)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
